---
title: "Pivot Lobsters"
author: "Julie Lowndes"
date: "1/27/2020"
output: html_document
---

Attaching our packages and reading in data

```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")
library(kableExtra) # install.packages("kableExtra")

# read in lobster data
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"),
                      skip = 4)
```

Viewing our data with skimr

```{r}
# these are equivalent
skimr::skim(lobsters) #explicitly stated from skimr
skim(lobsters) # not explicit
```

# Pivot Tables
In R, Pivot tables are: `group_by() %>% summarize()`
`n()` function will count observations
```{r}
# group by year
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())

# group by year and site
lobsters %>%
  group_by(year, site) %>%
  summarize(count_by_year_site = n())

# group by site and year
lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_site_year = n())

```

Summarize by multiple variables
And assign as an object. 
```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

Table formatting with `kableExtra::kable()`
```{r}
siteyear_summary %>%
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report. 

## Activity
1. build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm` and a function called `median()`
2. create a ggplot and save. For example:
```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm=TRUE))

ggplot(data = siteyear_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))  
```
3. knit and sync this all to github

```{r}
## attachtidyverse to demo tidyr::separate
library(tidyverse)

## look at the "smiths" data set 
smiths

## separate!
smiths %>%
  tidyr::separate(subject, into = c("first_name", "last_name"))


```


