---
title: "Making graphs with ggplot2"
author: "Allison Horst"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Get packages:
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(here)
library(readxl)
```

# California National Parks Data

```{r, message = FALSE}
ca_np <- read_csv(here("data","ca_np.csv"))

ci_np <- read_excel(here("data","ci_np.xlsx"))
```

### Make my own first graph! 
#### Here's a smaller subtitle

We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:

Command + Option/Alt + I
```{r}
# Trying ggplot2 with Channel Islands
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()

# Update the geom_ to geom_point(), geom_col(), geom_area()
```

Make a base layer for my CI plots:
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()
```

### Now let's start customizing!

- Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Activity: create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch), of a non-default size (size = #), and that are an updated color. 

```{r}
gg_base +
  geom_point(
    color = "dodgerblue",
    pch = 2,
    size = 4
  )
```

### Update graph aesthetics based on variables

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = year),
    pch = 10
  )

# Add an argument to update the color based on the year value
```

